### 2015-02-06 cav-2015-tex/bisimulation.tex

\section{Bisimilarity Proof}
\label{sec:bisim}
The theoretical result is the existence of a bisimulation between sets of states
states related by GSE ($p \rgse p^\prime$)
and states related by the algorithm in this paper ($q \rsym q^\prime$). The relations are on the universe of \emph{well-formed} states $S$, which have the
properties in \secref{sec:sh} with the constraint that
the states are \emph{feasible}: successors exist unless at
\cfgt{end}. Let $p \rgse p^\prime$ be a union over relations for GSE (see appendix):
$\rgse = \rgse^\mathit{A} \cup  \rgse^\mathit{A^\prime} \cup \rgse^\mathit{W} \cup \rgse^\mathit{W^\prime} \cup \rgse^\mathit{E} \cup \rgse^\mathit{E^\prime} \cup \rcom$, where \emph{A} is a field access after evaluating the expression for
the base reference $\lp \cfgt{*}\ \cfgt{\$}\ \cfgnt{f} \rightarrow
\cfgnt{k}\rp$, \emph{W} is a field write after evaluating the
expression for the right operand $\lp \cfgnt{x}\ \cfgt{\$}\ \cfgnt{f}\ \cfgt{:=}\ \cfgnt{*} \rightarrow
\cfgnt{k}\rp$, \emph{E} is a reference compare after evaluating the
left and right operands $\lp \cfgnt{v}\ \cfgt{=}\ \cfgnt{*}
\rightarrow \cfgnt{k}\rp$, the prime symbol indicates a null reference
in the operation or a false outcome, and \emph{\com} is everything
else in the language. Any state relation, say $\rightarrow_x$, is
trivially extended to sets of states as
$$
P \hookrightarrow_x P^\prime \Longleftrightarrow \forall p \in P\ (\forall p^\prime\ (p \rightarrow_x p^\prime \Rightarrow p^\prime \in P^\prime))
$$
Let $\hookrightarrow_\gse$ be the extension of $\rgse$ to sets of
states. From this extension, a new meta transition relation is defined
over sets of states as
$$
\rsgse = \hookrightarrow_\gse^\mathit{A}
\cup \hookrightarrow_\gse^\mathit{A^\prime} \cup \hookrightarrow_\gse^\mathit{W} \cup
\hookrightarrow_\gse^\mathit{W^\prime} \cup \hookrightarrow_\gse^\mathit{E} \cup \hookrightarrow_\gse^\mathit{E^\prime}
\cup \hookrightarrow_\com
$$
The relation captures the notion of splitting groups of heaps at
certain operations. For example, suppose we have a set $P$ containing a single state $p$
with all references uninitialized. If $p$ is a field access state, it has two
potential successors in GSE: a non-null reference and a null reference. 
Thus, the $\rsgse$ relation has two successors and divides $P$ into the two outcomes.

>>>>

\section{Bisimilarity Proof}
\label{sec:bisim}

%This section establishes the correctness of
%the~\symtxt{}. Demonstrating correctness is important because many
%summary heap techniques are correct only when certain, often
%unstated, assumptions are made. To show the absence of such
%assumptions, it is necessary to demonstrate soundness and completess
%with respect to all properties provable by symbolic execution. The
%proof relies on the observation that a state in symbolic execution
%can be thought of as representing a set of possible concrete
%states. Similarly, a state in the~\symtxt{} machine represents a set
%of possible~\gsetxt{} states. The goal is to show that states in
%the~\symtxt{} machine behave exactly like the sets of~\gsetxt{}
%states that they represent.

The bisimulation proof in this section establishes the soundness and
completeness of the symbolic heap approach with respect to Generalized
Symbolic Execution (GSE). Intuitively, any properties proven with GSE
can also be proved using the symbolic heap algorithm.  GSE concretizes
the possible heap configurations along with various aliasing
relationships when an uninitialized field is dereferenced. The rules
for initializing and updating the fields in the Javalite semantics for
GSE are shown in the~\appref{app:gse}. The proof demonstrates a bisimulation
relation between the states generated by GSE and the states generated
by the symbolic heap algorithm. The proof only reasons about
well-formed states, which have deterministic, type-consistent heaps,
and feasible environments, expressions, and continuations. Well-formed
states always have successors unless at the end of a program
execution. It is trivial to show that the set of well-formed states is
closed under the transition relation.

The main theoretical result establishes the existence of a
bisimulation between sets of states states related by GSE ($p \rgse
p^\prime$) and states related by the algorithm in this paper ($q \rsym
q^\prime$). The relations are on the universe of \emph{well-formed}
states $S$, which have the properties in \secref{sec:sh} with the
constraint that the states are \emph{feasible}: successors exist
unless at \cfgt{end}. Let $p \rgse p^\prime$ be a union over relations
for GSE (see appendix): $\rgse = \rgse^\mathit{A} \cup
\rgse^\mathit{A^\prime} \cup \rgse^\mathit{W} \cup
\rgse^\mathit{W^\prime} \cup \rgse^\mathit{E} \cup
\rgse^\mathit{E^\prime} \cup \rcom$, where \emph{A} is a field access
after evaluating the expression for the base reference $\lp
\cfgt{*}\ \cfgt{\$}\ \cfgnt{f} \rightarrow \cfgnt{k}\rp$, \emph{W} is
a field write after evaluating the expression for the right operand
$\lp \cfgnt{x}\ \cfgt{\$}\ \cfgnt{f}\ \cfgt{:=}\ \cfgnt{*} \rightarrow
\cfgnt{k}\rp$, \emph{E} is a reference compare after evaluating the
left and right operands $\lp \cfgnt{v}\ \cfgt{=}\ \cfgnt{*}
\rightarrow \cfgnt{k}\rp$, the prime symbol indicates a null reference
in the operation or a false outcome, and \emph{\com} is everything
else in the language. Any state relation, say $\rightarrow_x$, is
trivially extended to sets of states as
$$
P \hookrightarrow_x P^\prime \Longleftrightarrow \forall p \in P\ (\forall p^\prime\ (p \rightarrow_x p^\prime \Rightarrow p^\prime \in P^\prime))
$$
Let $\hookrightarrow_\gse$ be the extension of $\rgse$ to sets of
states. From this extension, a new meta transition relation is defined
over sets of states as
$$
\rsgse = \hookrightarrow_\gse^\mathit{A}
\cup \hookrightarrow_\gse^\mathit{A^\prime} \cup \hookrightarrow_\gse^\mathit{W} \cup
\hookrightarrow_\gse^\mathit{W^\prime} \cup \hookrightarrow_\gse^\mathit{E} \cup \hookrightarrow_\gse^\mathit{E^\prime}
\cup \hookrightarrow_\com
$$
The relation captures the notion of splitting groups of heaps at
certain operations. For example, suppose we have a set $P$ containing a single state $p$
with all references uninitialized. If $p$ is a field access state, it has two
potential successors in GSE: a non-null reference and a null reference. 
Thus, the $\rsgse$ relation has two successors and divides $P$ into the two outcomes.

### 2015-02-06 cav-2015-tex/lazy.tex

\section{Generalized Symbolic Execution}

The specialization of Javalite to generalized symbolic execution with
lazy initialization using the summary heap is in \figref{fig:lazyInit}
for the initialization and \figref{fig:lazy} for how the
initialization is used. Specifically, the field-access rule in
\figref{fig:lazy} uses initialization, $\rightarrow_I^*$, to be sure
the accessed field is instantiated. Initialization in generalized
symbolic execution never happens for more than one object on any use
of $\rightarrow_I^*$: $\Lambda$ is either empty or contains exactly one
location.

>>>>

\section{Generalized Symbolic Execution}
\label{app:gse}
The specialization of Javalite to generalized symbolic execution with
lazy initialization using the summary heap is in \figref{fig:lazyInit}
for the initialization and \figref{fig:lazy} for how the
initialization is used. Specifically, the field-access rule in
\figref{fig:lazy} uses initialization, $\rightarrow_I^*$, to be sure
the accessed field is instantiated. Initialization in generalized
symbolic execution never happens for more than one object on any use
of $\rightarrow_I^*$: $\Lambda$ is either empty or contains exactly one
location.
